---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
set.seed(1)
```

# qlcheckr

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/qlcheckr)](https://CRAN.R-project.org/package=qlcheckr)
[![R-CMD-check](https://github.com/learnr-academy/qlcheckr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/learnr-academy/qlcheckr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`qlcheckr` is an R package designed to analyse parsed user code by searching for specific patterns within its Abstract Syntax Tree (AST). The package is ideal for interactive code evaluation in Quarto Live environments, enabling precise checks for function usage, arguments, and expressions.

The `qlcheckr` package is designed with minimal dependencies to ensure a lightweight footprint, making it an ideal choice for integration with [Quarto Live](https://github.com/r-wasm/quarto-live) environments. By avoiding external package dependencies, `qlcheckr` significantly reduces file size and load times. The package design prioritises simplicity and speed, providing helpful tools without unnecessary bloat.

## Features

- **Function Usage Detection**: Verify if a specific function is used in the code.
- **Argument Matching**: Identify the presence of specific arguments in function calls.
- **Expression Searching**: Detect custom expressions within the code.
- **Flexible Pattern Matching**: Combine searches for functions, arguments, and expressions.

## Installation

To install the development version of `qlcheckr` from GitHub, use the `remotes` package:

```r
# Install remotes if not already installed
install.packages("remotes")

# Install qlcheckr
remotes::install_github("learnr-academy/qlcheckr")
```

## Usage

### Basic Examples

The `search_ast` function allows you to search for specific patterns in parsed R code.

#### Search for a Function Name

```{r}
library(qlcheckr)

code <- quote(mean(x))
search_ast(code, .fn = "mean")
```

#### Search for an Argument

```{r}
code <- quote(mean(x, na.rm = TRUE))
search_ast(code, na.rm = TRUE)
search_ast(code, na.rm = FALSE)
```

#### Search for an Expression

```{r}
code <- quote(mean(x + y))
search_ast(code, .expr = x + y)
```


### Using `exists_in`

The `exists_in()` function checks whether elements in a list satisfy specific conditions. It is particularly useful when working with multiple outputs to pinpoint a result or identify errors.

The `.evaluate_result` object available in the check chunk of quarto-live exercises contains information about results, errors, warnings, messages and more. A manually constructed `.evaluate_result` is shown here for testing purposes:

```{r}
.evaluate_result <- evaluate::evaluate(
'print(rnorm(10))
sample$x
log(-1)
message("Hello world")',
output_handler = ql_output_handler)
```


<!-- TODO: Create dummy `.evaluate_results` and directly use `ql_results()`, `ql_errors()`, ... -->

#### Find a Matching Result

Suppose you are processing a list of numeric outputs and want to find if any value is greater than a threshold:

```{r}
ql_results()

exists_in(ql_results(), ~ any(. > 0.5))
```

This helps quickly confirm if any result (among many outputs) passes your test.

#### Detecting specific errors

Consider checking code that might generate errors, which can be obtained in Quarto Live check environments with the `ql_errors()` helper. `exists_in()` can again be used to search for specific errors.

```{r}
ql_errors()

exists_in(ql_errors(), grepl, pattern = "not subsettable")
exists_in(ql_errors(), grepl, pattern = "non-numeric argument")
```

This is valuable for identifying and debugging specific errors among many possible results.

It is also possible to search warnings and messages with similar helpers:

```{r}
ql_warnings()
ql_messages()
```



## Contributing

We welcome contributions to `qlcheckr`! If you encounter any issues or have suggestions for improvement, please open an issue or submit a pull request on [GitHub](https://github.com/learnr-academy/qlcheckr).

## License

`qlcheckr` is licensed under the MIT License. See the [LICENSE](LICENSE) file for details.

